---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages and cleaned data
```{r}
#load needed packages. make sure they are installed.
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(dotwhisker) #visualizing regression results

#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")

#load cleaned data. 
mydata <- readRDS(data_location)
```

# fit linear model to contiuous outcome (BodyTemp) - only main predictor (RunnyNose)
```{r}
ggplot(mydata, aes(x = RunnyNose, y = BodyTemp)) + geom_point() + geom_smooth(method = lm)

#linear regression model specification
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training model
lm_fit <- lm_mod %>% fit(BodyTemp ~ RunnyNose, data = mydata)
#view summary of fit
tidy(lm_fit)
```
Not sure if this is working the way I want it to, will look into further later. 


# fit linear model to continuous outcome (BodyTemp) using all predictors
```{r}

```

# compare model results for main predictor (RunnyNose) vs all predictors
```{r}

```

# fit logistic model to categorical outcome (Nausea) using main predictor of interest (RunnyNose)
```{r}


```

# fits logistic model to categorical outcome (Nausea) with all predictors of interest
```{r}

```

# compare model results with main predictor (RunnyNose) vs all predictors
```{r}
```


```{r}
```

