---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages and cleaned data
```{r}
#load needed packages. make sure they are installed.
library(ggplot2) #for plotting
library(broom) #for cleaning up output from lm()
library(here) #for data loading/saving
library(tidymodels) #for modeling
library(dotwhisker) #visualizing regression results

#path to data
#note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")

#load cleaned data. 
mydata <- readRDS(data_location)
```

# fit linear model to contiuous outcome (BodyTemp) - only main predictor (RunnyNose)
```{r}
ggplot(mydata, aes(x = RunnyNose, y = BodyTemp)) + geom_point() + geom_smooth(method = lm)

#linear regression model specification set up
lm_mod <- linear_reg() %>% set_engine("lm")
#estimating/training linear model
lm_fit1 <- lm_mod %>% fit(BodyTemp ~ RunnyNose, data = mydata)
#view summary of fit
tidy(lm_fit1)
#save as table
lm_fit1_table <- tidy(lm_fit1)
```

# fit linear model to continuous outcome (BodyTemp) using all predictors
```{r}
# running linear model 2
lm_fit2 <- lm_mod %>% fit(BodyTemp ~ ., data = mydata)
#view summary of fit
tidy(lm_fit2)
#save as table
lm_fit2_table <- tidy(lm_fit2)
```

# compare model results for main predictor (RunnyNose) vs all predictors
```{r}
#display model results
glance(lm_fit1)
glance(lm_fit2)

#use ANOVA to compare model with one predictor vs model with all predictors
lm_comp <- anova(lm_fit1$fit, lm_fit2$fit, test = "Chisq")
lm_comp
```

The ANOVA test tells us that we should use the second linear model, which includes all predictors. lm_fit2 is a statistically significant model

# fit logistic model to categorical outcome (Nausea) using main predictor of interest (RunnyNose)
```{r}
#logistic model set up
log_mod <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

#run logistic model for RunnyNose predictor
log_fit1 <- log_mod %>% fit(Nausea ~ RunnyNose, data = mydata)

```

# fits logistic model to categorical outcome (Nausea) with all predictors of interest
```{r}
#run logistic model with all predictors
log_fit2 <- log_mod %>% fit(Nausea ~ ., data = mydata)
```

# compare model results with main predictor (RunnyNose) vs all predictors
```{r}
#display logistic model results
glance(log_fit1)
glance(log_fit2)

#compare model results using ANOVA
log_comp <- anova(log_fit1$fit, log_fit2$fit, test = "Chisq")
log_comp
```

This ANOVA test tells us that the more complex model, log_fit2, which uses all predictors, is a more suitable model (having a significant p-value). 

#saving results
```{r}
saveRDS(lm_comp, file = here("results", "LinearComp.rds"))
saveRDS(log_comp, file = here("results", "LogisticComp.rds"))
```

#testing out KNN
```{r}
install.packages("kknn") #install KNN package
library(kknn) #load KNN package
#set up K-nearest neighbors model fit
knn_mod <-  nearest_neighbor() %>% 
  set_mode("regression")
#fit model for BodyTemp and all predictors
knn_fit1 <- knn_mod %>% fit(BodyTemp ~ ., data = mydata)
#view model
knn_fit1 

```
Okay, now I have created the model but I am still working on how to intepret it's accuracy/success. We'll get back to this in a bit